---
title: "Data Exploration Project - Data Cleaning"
author: "Elias Kazemi"
date: "2/16/2022"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Packages

```{r, message = FALSE}
library(purrr)
library(dplyr)
library(tidyverse)
library(lubridate)
```

# Data Joining and Cleaning

First, we create the dataframe with merged data from all of the trends files. the trends_up_to_inter_2.csv file's schid column has a file type of character, so we convert this to integer to match the rest of the files and be able to bind rows:

```{r}
flist <- list.files(pattern = 'trend')
flist <- flist %>% 
  map(read.csv)

flist[[2]]$schid <- as.integer(flist[[2]]$schid)

df <- bind_rows(flist)
```
Some colleges have names that are the same. Since these are difficult to separate in our analysis, we are only keeping the first occurrence of a college name in our analysis:

```{r}
id_name_link <- read.csv('id_name_link.csv')
id_name_link <- id_name_link %>% 
  group_by(schname) %>% 
  mutate(N = n()) %>% 
  filter(N == 1)
```

We can join the trends data with the id_name_link data by schname:

```{r}
dfnamelink <- df %>% 
  left_join(id_name_link, by = 'schname')
```

And then join the scorecard data to this joined data frame by OPEID. Note that we have to change the column name to uppercase so they join together.

```{r}
scorecard <- read.csv('Most+Recent+Cohorts+(Scorecard+Elements).csv')

joineddf <- dfnamelink %>% 
  rename(OPEID = opeid) %>% 
  left_join(scorecard, by = 'OPEID')
```

We then have our final dataframe with a new variable index_std to return standardized results of our index variable so that the outcome is relative to each individual school's results.

```{r}
finaldf <- joineddf %>%
  group_by(schname, keyword) %>%
  mutate(index_std = (index - mean(index,na.rm = TRUE))/sd(index, na.rm = TRUE))
```

We also want to create a singular date variable to see the effect of the release of the scorecard:

```{r}
finaldf <- finaldf %>%
  mutate(date = str_sub(monthorweek, 1, 10)) %>%
  mutate(date = ymd(date)) %>%
  mutate(AfterScorecardRelease = date > ymd('2015-09-01'))
```

We can then drop any irrelevant variables by selecting the relevant ones for our analysis. We are keeping any variables we think may be useful in our analysis later on:

```{r}
finaldf <- finaldf %>% 
  select(c(keyword, AfterScorecardRelease, PREDDEG, md_earn_wne_p10.REPORTED.EARNINGS, INSTNM, CITY, STABBR, CONTROL, CURROPER, index_std, date))
```

# Data Transformation

We want our scorecard value to be binary and our earnings variable to be numeric, making those quick adjustments:

```{r}
finaldf$AfterScorecardRelease <- as.factor(finaldf$AfterScorecardRelease)
finaldf$md_earn_wne_p10.REPORTED.EARNINGS <- as.numeric(finaldf$md_earn_wne_p10.REPORTED.EARNINGS)
```


# Writing Data to New Files


```{r}
write.csv(finaldf, 'cleandata.csv', row.names = TRUE)
```

